---
title: "CellChat_celltype_overall_detail"
author: "Mitsuhiro Yuasa, Haryoon Kim"
date: "2024/04/10"
output: html_document
editor_options: 
chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

##########
knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
##########
```

```{r load libraries}
library(tidyverse)
library(gsubfn)
library(readxl)
library(RColorBrewer)
library(viridis)
library(ggsci)
library(grid)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(exact2x2)
library(lawstat)
library(ggbeeswarm)
library(scales)
theme_set(theme_cowplot())
library(Seurat)
library(Matrix)
library(broom)
library(ComplexHeatmap)
library(circlize)
library(CellChat)
library(patchwork)
```

```{r source}
source("R/utilities.R")
source("ref/cluster_annotation/cluster_annotation_level_220209.R")
```

```{r make tst and directory}
tst <- "2023-08-29"

# pdf
outdir_pdf <- str_glue("result/REVISION/{tst}/4_Data_filt3/14_CellInteraction/02_CellChat_celltype_overall_detail/pdf")
dir.create(outdir_pdf, recursive = TRUE, showWarnings = FALSE)

# table
outdir_table <- str_glue("result/REVISION/{tst}/4_Data_filt3/14_CellInteraction/02_CellChat_celltype_overall_detail/table")
dir.create(outdir_table, recursive = TRUE, showWarnings = FALSE)

# Rds
outdir_Robj <- str_glue("result/REVISION/{tst}/4_Data_filt3/14_CellInteraction/02_CellChat_celltype_overall_detail/rds")
dir.create(outdir_Robj, recursive = TRUE, showWarnings = FALSE)
```

# 1. Preparation
```{r load Seurat obj}
mouseNK_data <- readRDS("result/table/Robj/aggr/mouseNK_data_RNAdim45dim011_afterExclude_v1_other3660SP.rds")
```

```{r load metadata}
Annotation_metadata_filt1_df <- read_tsv("result/REVISION/2023-08-29/4_Data_filt3/3_reclutser/9_Reannotation/06_integrate_subclusters/table/metadata_mouseNK_data_celltype_overall.txt")

Annotation_metadata_filt1_df$cellbarcode2 %>% length()
```

```{r combine metadata}
Annotation_metadata_filt1_df %>% colnames()
mouseNK_data@meta.data %>% colnames()

setdiff(colnames(Annotation_metadata_filt1_df), colnames(mouseNK_data@meta.data))
unique(c("cellbarcode2", setdiff(colnames(Annotation_metadata_filt1_df), colnames(mouseNK_data@meta.data))))

annotation_df <- Annotation_metadata_filt1_df %>% 
  dplyr::select(unique(c("cellbarcode2", 
                         setdiff(colnames(Annotation_metadata_filt1_df), 
                                 colnames(mouseNK_data@meta.data))))) %>% 
  as.data.frame()
rownames(annotation_df) <- annotation_df$cellbarcode2
annotation_df %>% head()

mouseNK_data <- AddMetaData(object = mouseNK_data, metadata = annotation_df)
```

```{r add diseasetype}
mouseNK_data@meta.data <- mouseNK_data@meta.data %>% 
  mutate(
    disease_type_edit = case_when(
      disease_type %in% c("mNK_SG", "mNK_Tu") ~ "SG/Other",
      disease_type == "mNK_SP"                ~ "SP"
    ),
    disease_type2_edit = case_when(
      disease_type2 %in% c("CreN_pre", "CreP_pre") ~ "Pre-tumor",
      disease_type2 == "CreP_Tumor"                ~ "Trp53-/- tumor",
      disease_type2 == "CreLMP1P_Tumor"            ~ "Trp53-/- LMP1+ tumor"
    ),
    disease_type3 = case_when(
      disease_type == "mNK_SG" & disease_type2 == "CreN_pre"       ~ "SG_CreN_pre",
      disease_type == "mNK_SG" & disease_type2 == "CreP_pre"       ~ "SG_CreP_pre",
      disease_type == "mNK_SP" & disease_type2 == "CreN_pre"       ~ "SP_CreN_pre",
      disease_type == "mNK_SP" & disease_type2 == "CreP_pre"       ~ "SP_CreP_pre",
      disease_type == "mNK_SG" & disease_type2 == "CreP_Tumor"     ~ "SG_CreP_LMPn_tumor",
      # disease_type == "mNK_SG" & disease_type2 == "CreLMP1P_Tumor" ~ "SG_CreP_LMPp_tumor",
      disease_type == "mNK_SP" & disease_type2 == "CreP_Tumor"     ~ "SP_CreP_LMPn_tumor",
      disease_type == "mNK_SP" & disease_type2 == "CreLMP1P_Tumor" ~ "SP_CreP_LMPp_tumor",
      disease_type == "mNK_Tu" & disease_type2 == "CreLMP1P_Tumor" ~ "Tu_CreP_LMPp_tumor"
    )
  ) %>% 
  mutate(
    disease_type = fct_relevel(disease_type, c("mNK_SG", "mNK_Tu", "mNK_SP")),
    disease_type_edit = fct_relevel(disease_type_edit, c("SG/Other", "SP")),
    disease_type2 = fct_relevel(disease_type2, c("CreN_pre", "CreP_pre", "CreP_Tumor", "CreLMP1P_Tumor")),
    disease_type2_edit = fct_relevel(disease_type2_edit, c("Pre-tumor", "Trp53-/- tumor", "Trp53-/- LMP1+ tumor")),
    celltype_overall = fct_relevel(celltype_overall, c("CD4T", "CD8T", "B", 
                                                       "Myeloid", "NK", "Other", 
                                                       "Tumor_LMPn", "Tumor_LMPp")),
    celltype_overall_detail = fct_relevel(celltype_overall_detail, 
                                          c("CD4T_N", "CD4T_EM", "TREG",
                                            "CD8T_N", "CD8T_CM", "CD8T_EM", "CD8T_RM", "CD8T_EX",
                                            "B_Immature", "B_Mature", "Plasmablast",
                                            "Mono", "Macro", "cDC1", "cDC2", "pDC", "mregDC", "MDSC",
                                            "NK", "Other", "Tumor_LMPn", "Tumor_LMPp")),
  )
```

```{r subset}
# analyze only tumor samples
Idents(mouseNK_data) <- mouseNK_data$disease_type2_edit
mouseNK_data <- subset(x = mouseNK_data, idents = "Pre-tumor", invert = TRUE)

Idents(mouseNK_data) %>% unique()

gc()

# exclude "Other"
Idents(mouseNK_data) <- mouseNK_data$celltype_overall_detail
mouseNK_data <- subset(x = mouseNK_data, idents = "Other", invert = TRUE)

Idents(mouseNK_data) %>% unique()

gc()

levels_fixed <- setdiff(levels(mouseNK_data@meta.data$celltype_overall_detail), "Other")
mouseNK_data@meta.data$celltype_overall_detail <- factor(mouseNK_data@meta.data$celltype_overall_detail, levels_fixed)
```

# Part I: Data input & processing and initialization of CellChat object

## Load data
```{r}
data.input <- mouseNK_data@assays$SCT@data
meta <- mouseNK_data@meta.data
cell.use <- rownames(meta)[meta$celltype_overall_detail %in% levels_fixed]

data.input <- data.input[, cell.use]
meta <- meta[cell.use, ]
```

## Create a CellChat object
```{r}
cellchat <- createCellChat(object = data.input, meta = meta, group.by = "celltype_overall_detail")
```

## Add cell information into meta slot of the object (Optional)
```{r}
cellchat <- addMeta(cellchat, meta = meta)
cellchat <- setIdent(cellchat, ident.use = "celltype_overall_detail")

levels(cellchat@idents)
unique(cellchat@idents)

groupSize <- as.numeric(table(cellchat@idents))
names(groupSize) <- levels(cellchat@idents)
groupSize
```

## Set the ligand-receptor interaction database
```{r}
CellChatDB <- CellChatDB.mouse
CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling")
cellchat@DB <- CellChatDB.use
```

## Preprocessing the expression data for cell-cell communication analysis
```{r}
cellchat <- subsetData(cellchat)

future::plan("multisession", workers = 4)
options(future.globals.maxSize = 50000 * 1024^2)

cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
```

# Part II: Inference of cell-cell communication network

## Compute the communication probability and infer cellular communication network
```{r}
cellchat@idents <- factor(cellchat@idents, levels = levels_fixed)

cellchat <- computeCommunProb(cellchat)
cellchat <- filterCommunication(cellchat, min.cells = 10)
```

## Extract the inferred cellular communication network as a data frame
```{r}
df.net <- subsetCommunication(cellchat)

# for supplementary table
df.net %>% 
  dplyr::select(-annotation, -evidence) %>% 
  dplyr::rename("Source" = source,
                "Target" = target,
                "Ligand" = ligand,
                "Receptor" = receptor,
                "Communication probability" = prob,
                "P-value" = pval,
                "Interaction name" = interaction_name_2,
                "Pathway name" = pathway_name) %>% 
  dplyr::select(`Interaction name`, `Pathway name`, 
                Source, Target, Ligand, Receptor, 
                `Communication probability`, `P-value`) %>% 
  write_tsv(str_glue("{outdir_table}/df_inferred_cellular_communication_network_LR_for_suppl.txt"))
```

## Infer the cell-cell communication at a signaling pathway level
```{r}
cellchat <- computeCommunProbPathway(cellchat)
df.netP <- subsetCommunication(cellchat, slot.name = "netP")
```

## Calculate the aggregated cell-cell communication network
```{r}
cellchat <- aggregateNet(cellchat)
```

# Part III: Visualization of cell-cell communication network

```{r}
pathways.show.all <- cellchat@netP$pathways
```

```{r}
for (i in 1:length(pathways.show.all)) {
  # i <- 10
  this_pathway <- pathways.show.all[i]
  pairLR.this <- extractEnrichedLR(cellchat,
                                   signaling = this_pathway,
                                   geneLR.return = FALSE)
  for (k in 1:nrow(pairLR.this)) {
    LR.show <- pairLR.this[k,]

    # Chord diagram
    pdf(str_glue("{outdir_pdf}/Fig5f_netVisual_individualchord_{this_pathway}_{LR.show}.pdf"), width = 15, height = 16)
    netVisual_individual(cellchat,
                         signaling = this_pathway,
                         pairLR.use = LR.show,
                         layout = "chord")
    dev.off()
  }
}
```

# Part IV: Systems analysis of cell-cell communication network

## Identify signaling roles (e.g., dominant senders, receivers) of cell groups as well as the major contributing signaling

### Compute the network centrality scores
```{r}
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")
```

### Identify signals contributing most to outgoing or incoming signaling of certain cell groups
```{r}
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, 
                                         width = 20,
                                         height = 20,
                                         font.size = 12,
                                         font.size.title = 12,
                                         pattern = "outgoing")
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, 
                                         width = 20,
                                         height = 20,
                                         font.size = 12,
                                         font.size.title = 12,
                                         pattern = "incoming")

pdf(str_glue("{outdir_pdf}/FigS6d_netAnalysis_signalingRole_heatmap_both.pdf"), width = 35, height = 20)
ht1 + ht2
dev.off()

# save the source of heatmap
mat_outgoing <- matrix(0, 
                       nrow = length(cellchat@netP$pathways), 
                       ncol = length(levels(cellchat@idents)))
rownames(mat_outgoing) <- cellchat@netP$pathways
colnames(mat_outgoing) <- levels(cellchat@idents)
mat_incoming <- matrix(0, 
                       nrow = length(cellchat@netP$pathways), 
                       ncol = length(levels(cellchat@idents)))
rownames(mat_incoming) <- cellchat@netP$pathways
colnames(mat_incoming) <- levels(cellchat@idents)

for (i in 1:length(cellchat@netP$centr)) {
  # i <- 10
  this_pathway <- pathways.show.all[i]
  mat_outgoing[this_pathway,] <- cellchat@netP$centr[[this_pathway]]$outdeg
  mat_incoming[this_pathway,] <- cellchat@netP$centr[[this_pathway]]$indeg
}

mat_outgoing %>% 
  as.data.frame() %>% 
  rownames_to_column("Pathway") %>% 
  as_tibble() %>% 
  write_tsv(str_glue("{outdir_table}/mat_netAnalysis_signalingRole_outgoing.txt"))

mat_incoming %>% 
  as.data.frame() %>% 
  rownames_to_column("Pathway") %>% 
  as_tibble() %>% 
  write_tsv(str_glue("{outdir_table}/mat_netAnalysis_signalingRole_incoming.txt"))

# save the scaled source
ht1@matrix %>% 
  as.data.frame() %>% 
  rownames_to_column("Pathway") %>% 
  as_tibble() %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  write_tsv(str_glue("{outdir_table}/mat_netAnalysis_signalingRole_outgoing_scaled.txt"))

ht2@matrix %>% 
  as.data.frame() %>% 
  rownames_to_column("Pathway") %>% 
  as_tibble() %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  write_tsv(str_glue("{outdir_table}/mat_netAnalysis_signalingRole_incoming_scaled.txt"))
```

# Part V: Save the CellChat object

```{r}
saveRDS(cellchat, file = str_glue("{outdir_Robj}/mouseNK_data_cellchat_Tumor.rds"))
```

```{r sessioninfo}
sessionInfo()
```
